Mixture Models for Single-Cell Assays with Application to Vaccine Studies

### 2013-05-23 MIMOSA.tex
Update fluidigm

\subsection{Single-cell gene expression}
We applied the MIMOSA model to a Fluidigm single-cell gene expression data set. We used the two-sided MIMOSA model because genes could be regulated upward or downward upon stimulation. In order to detect stimulation specific changes of expression, we fit our model to each gene within each stimulation. The results presented in Figure~\ref{fig:fluidigm} show that MIMOSA identifies stimulation-specific differences in the proportions of cells expressing each gene while preserving inter-subject variability (Figure~\ref{fig:fluidigm} A,B). These patterns are evident in the  posterior probabilities (Figure~\ref{fig:fluidigm} A) and preserved in the posterior estimates of the differences of proportions (Figure~\ref{fig:fluidigm} B). A similar analysis using a two-sided Fisher's exact test and clustering the signed FDR adjust p-values (Figure~\ref{fig:fluidigm} C) does not reveal any stimulation-specific patterns. For an FDR of 10\%, Fisher's exact test identified 47 significant genes, while MIMOSA identified 50 significant genes. Both methods identified 39 genes in common.

>>>>

\subsection{Single-cell gene expression}
We applied the MIMOSA model to a Fluidigm single-cell gene expression data set. We used the two-sided MIMOSA model because genes could be regulated upward or downward upon stimulation. In order to detect stimulation specific changes of expression, we fit our model to each gene within each stimulation. The results presented in Figure~\ref{fig:fluidigm} show that MIMOSA identifies stimulation-specific differences in the proportions of cells expressing each gene while preserving inter-subject variability (Figure~\ref{fig:fluidigm} A,B). These patterns are evident in the  posterior probabilities (Figure~\ref{fig:fluidigm} A) and preserved in the posterior estimates of the differences of proportions (Figure~\ref{fig:fluidigm} B). A similar analysis using a two-sided Fisher's exact test and clustering the signed FDR adjust p-values (Figure~\ref{fig:fluidigm} C) does not reveal any stimulation-specific patterns. For an FDR of 10\%, Fisher's exact test identified 13 significant genes, while MIMOSA identified 46 significant genes. The 13 genes identified by Fisher's method were a subset of those identified by MIMOSA.
